Foreign Exchange Speculation: An Event Study

### 2017-12-08 MDPI/MDPI_template/FinStudies.tex
Updated the files for the publication.

\begin{document}
\section{Introduction}
If speculation is based on information it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. \citet{KeynesHedge} and \citet{HicksHedge} for example emphasised the role that speculation plays in facilitating financial market transactions. Speculators offer a service for a fee that is determined by the return on their activity.  While speculators may be trading on information or noise, informed speculators will tend to drive out the uninformed noise-traders according to \citet{FriedmanPositive}.   Nonetheless, the later Keynes regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'\citet[p. 101]{Keynes1936} and \citet{Delong1990noise} provides a model of cases where speculative noise provides a constrain on informed traders benefiting from their knowledge.

This paper evaluates the relative importance of these two concepts of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to subsequent movement of foreign exchange prices.\footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.  The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).}.  If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes.

@@

\section{Informed and uninformed speculation}
Speculation plays an important part in several areas of financial economics.  However, there is no clear, unambiguous theory of speculation.  Speculation is frequently associated with the term \emph{trading}, but trading is broader as it encompasses market-making and financial investment. Speculation and trading can be categorised as being informed or uninformed. Those trading without information are frequently called \emph{noise-traders} or \emph{liquidity traders}.  However, the terms are not fixed and  \citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.  See \citet{Ramiah201589} for a recent review of the noise-trader literature.

The comparison of informed and uninformed traders is a central component of microstructure research.  The focus of the \citet{Kyle1985Continuous} model of market-making is the interaction of orders from the informed and the uninformed.  On the assumption that uniformed orders arrive in a random fashion, the weight of orders reveals the direction of private information and should led to adjustment of the bid-ask spread as part of the process of price discovery.  Similarly,  \citet{glosten1985bid} describes a model where the bid-ask spread is a function of adverse selection, being caused by the risk that orders come from informed traders.

@@

\section{Measuring speculation}
This study uses a unique series of option\emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.  The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.  The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.  Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put.    If the exchange rate is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.  Any deviation from this equivalence is an indication that a divergence from the random walk with drift is implied in the market price of options.\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.  The dealer pays 0.15\%.  Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\%. See \citet{Global} for a technical review. }  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity is an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls and as an indication of market sentiment.  The larger the absolute level of the risk-reversal skew, the greater the relative price and, if the deviation is based on information, subsequent price action should be zero on average; if the deviation is based on noise, this movement should be reversed.  See \citet{Yan2011} and \citet{FengZhangFriesen} for some analysis of implied volatility in the equity market.

@@

The effect of extreme speculation on future returns is analysed with an event study.  The events here are the extremes of speculative sentiment and speculative positions.  See \citet{Dolly1933} for the original version of the event study method and \citet{FamaFisherJensenRoll} for an overview of the method used in this study.

@@

Does knowledge about the extreme at event time $t = 0$ provide information about the future returns?  If this speculation is noise, informed traders, such as real money or long-term fundamental accounts, should take advantage of the space that has been created between the price and the fundamental value and a reversal should be seen.  Whether this is the case is recorded in columns 5, 8, 11 and 14 of Table \ref{tabref:RR1}. A reversal is indicated by a significant negative reading for highs or positive reading for lows.   Looking, for example, at the four-day window in column 8, it is clear that the extreme risk reversal skew does not provide any information about future returns.  Of the eighteen cases using the $5^{th}$ and $95^{th}$ percentiles to identify extremely negative and extremely positive sentiment respectively, ten show a continuation (rather than a reversal) that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal - EURCHF in each case.  For the more extreme event when the $99^{th}$ percentile is used, there is even less clarity.  Just over half the cases with the four-day window show a reversal (8 of 18), but only 2 of these are statistically significant. Exchange rate prices are as likely to go up as down after an extreme and there does not appear to be any information about the future from knowledge of these extremes. Extremes seem to be followed by a random walk as would be the case if the extreme were the result of information being absorbed into the price.  
\begin{figure}
\graphicspath{{../Figures/}}
\centering
\includegraphics[scale=0.8]{RRCum16}
\caption{Event Study: Extreme (High $99{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is at or above the $99^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns. The event day is the day of the extreme risk reversal reading.  If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk. The solid line is the cumulative mean return from the start of the event window; the dashed lines show the 95\% confidence intervals constructed from 1000 random bootstrap samples of appropriate event window day.}
\label{fig:ES5}
\end{figure}

@@

\begin{figure}
\graphicspath{{../Figures/}}
\centering
\includegraphics[scale=0.8]{RRCum16}
\caption{Event Study: Extreme (High $99{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is at or above the $99^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns. The event day is the day of the extreme risk reversal reading.  If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk. The solid line is the cumulative mean return from the start of the event window; the dashed lines show the 95\% confidence intervals constructed from 1000 random bootstrap samples of appropriate event window day.}
\label{fig:ES1}
\end{figure}

@@

There is little evidence that these self-reported speculators are driving prices away from fundamentals with their activity. Sixty percent of the S1 cases still show a significant abnormal return in the direction of the extreme 4 weeks after the extreme has been reached; only thirty percent of the cases show a reversal and in only one of those do ninety five percent of the means calculated after re-sampling remain above zero.  For S2 there is even less clarity, with ninety percent of the cases showing a continuation and forty percent of them showing significant continuation.  The evidence here with speculative positions is the same as that of speculative sentiment.  Prices move in the direction of speculative activity but speculative extremes do not provide information about the future. The price action of the event study is consistent with speculation being informed rather than uninformed.

>>>>

\begin{document}
\section{Introduction}
If speculation is based on information it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. \citet{KeynesHedge} and \citet{HicksHedge} for example emphasise the role that speculation plays in facilitating financial market transactions. Speculators offer a service for a fee that is determined by the return on their activity.  While speculators may be trading on information or noise, informed speculators will tend to drive out the uninformed noise-traders according to \citet{FriedmanPositive}.   However, the later \citet[p. 101]{Keynes1936} regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd' and \citet{Delong1990noise} provides a model of cases where speculative noise provides a constrain on informed traders benefiting from their knowledge. Now speculation is an uninformed and destabilising force.

This paper evaluates the relative importance of these two concepts of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to subsequent movement of foreign exchange prices.\footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2016 estimated average daily turnover at \$5.1trn with average daily spot transactions of \$1.7trn See \citet{BISFX2016}.  The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).}  If speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes; if speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place.

@@

\section{Informed and uninformed speculation}
Speculation plays an important part in several areas of financial economics.  However, there is no clear, unambiguous theory of speculation.  Speculation is frequently associated with the term trading and the provision of liquidity, but trading is broader as it encompasses financial investment while liquidity provision is really market-making. Speculation and trading can be categorised as being informed or uninformed. Those trading without information are frequently called \emph{noise-traders} or \emph{liquidity traders}.  However, the terms are not fixed and  \citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.  See \citet{Ramiah201589} for a recent review of the noise-trader literature.

The comparison of informed and uninformed traders is a central component of microstructure research.  At the heart of the \citet{Kyle1985Continuous} model of market-making is the interaction of orders from the informed and the uninformed.  On the assumption that uniformed orders arrive in a random fashion, the weight of orders reveals the direction of private information and should encourage adjustment of the bid-ask spread as part of the process of price discovery.  \citet{glosten1985bid} presents a model where the bid-ask spread is a function of adverse selection, being caused by the risk that orders come from informed traders.

@@

\section{Measuring speculation}
This study uses a unique series of option \emph{risk-reversal skews} to measure of speculative sentiment\footnote{It appears that this is the only period for which the data currently exist.  The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.  The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.  Risk-reversal skew is the relative price of currency put and call options for the same strike price.}, as well as CFTC data on speculative positions. The risk-reversal skew is the relationship between a call option and the equivalent put.    If the exchange rate is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.  Any deviation from this equivalence is an indication that a divergence from the random walk with drift is implied in the market price of options.\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.  The dealer pays 0.15\%.  Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\%. See \citet{Global} for a technical review. }  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity strike is an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls and as an indication of market sentiment or related market positioning.  The larger the absolute level of the risk-reversal skew, the greater the relative price and, if the deviation is based on information, subsequent price action should be zero on average; if the deviation is based on noise, this movement should be reversed.  See \citet{Yan2011} and \citet{FengZhangFriesen} for some analysis of implied volatility in the equity market.

@@

An event study will analyse the performance of security of asset prices around an event.  In this case the event is the extreme of speculative sentiment or extreme speculative activity.  See \citet{Dolly1933} for the original version of the event study method and \citet{FamaFisherJensenRoll} for an overview of the method used in this study.

@@

Does knowledge about the extreme at event time $t = 0$ provide information about the future returns?  If this speculation is noise, informed traders, such as real money or long-term fundamental accounts, should take advantage of the space that has been created between the price and the fundamental value and a reversal should be seen.  Whether this is the case is recorded in columns 5, 8, 11 and 14 of Table \ref{tabref:RR1}. A reversal is indicated by a significant negative reading for highs or positive reading for lows.   Looking, for example, at the four-day window in column 8, it is clear that the extreme risk reversal skew does not provide any information about future returns.  Of the eighteen cases using the $5^{th}$ and $95^{th}$ percentiles to identify extremely negative and extremely positive sentiment respectively, ten show a continuation (rather than a reversal) that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal - EURCHF in each case.  For the more extreme event when the $99^{th}$ percentile is used, there is even less clarity.  Just over half the cases with the four-day window show a reversal (8 of 18), but only 2 of these are statistically significant. Exchange rate prices are as likely to go up as down after an extreme and there does not appear to be any information about the future from knowledge of these extremes. Extremes seem to be followed by a random walk as would be the case if the extreme were the result of information being absorbed into the price.  Figures \ref{fig:ES5} and \ref{fig:ES5} show details of a selection of the event studies.  Figure \ref{fig:ES5} is the extreme high ($99^{th}$ percentile) with a 16 day window. There is no evidence of a reversal in any of the currency units.   Figure \ref{fig:ES5} is the extreme low ($10^{th}$ percentile) with a 16 day event window.  Only the CAD shows any evidence of reversal.

\begin{figure}
\graphicspath{{../Figures/}}
\centering
\includegraphics[scale=0.8]{RRCum16}
\caption{Event Study: Extreme (High $99{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is at or above the $99^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns. The event day is the day of the extreme risk reversal reading.  If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk. The solid line is the cumulative mean return from the start of the event window; the dashed lines show the 95\% confidence intervals constructed from 1000 random bootstrap samples of appropriate event window day.}
\label{fig:ES5}
\end{figure}

@@

@@

There is little evidence that these self-reported speculators are driving prices away from fundamentals with their activity. Sixty percent of the S1 cases still show a significant abnormal return in the direction of the extreme 4 weeks after the extreme has been reached; only thirty percent of the cases show a reversal and in only one of those do ninety five percent of the means calculated after re-sampling remain above zero.  For S2 there is even less clarity, with ninety percent of the cases showing a continuation and forty percent of them showing significant continuation.  The evidence here with speculative positions is the same as that of speculative sentiment.  Prices move in the direction of speculative activity but speculative extremes do not provide information about the future. The price action of the event study is consistent with speculation being informed rather than uninformed.  Figures \ref{fig:ES3} and \ref{fig:ES4} show a selection of event studies that were carried out.  Figure \ref{fig:ES3} is the extreme low ($10^{th}$ percentile) with a 13 week window. There is no evidence of a reversal in any of the currency units.   Figure \ref{fig:ES4} is the extreme high ($90^{th}$ percentile) with a 13 week event window.  Only the GBP shows any evidence of reversal. It is not statistically significant.
